The binary liquid model for gravity-driven microchannel simulations

### 2010-10-06 Text/paper.tex
Added small introduction.

\begin{document}
\section{Introduction}
The Taylor flow which has analytical solution for low Capillary numbers
\cite{bretherton} is the attractive benchmark for the diffusion interface
models. The film width needed to be resolved is proportional to $Ca^{2/3}$
which is for small capillary numbers can reach numbers from $0.001$ to $0.01$.
When it comes to the resolution of the film thickness the diffusion interface
should be resolved in such way as to have negligible effect on physics in the
film. The mass transfer which exists in the small film drops the bubble driven
pressure to proportionality $Ca^{2/3}$ \cite{kreutzer-pressure-drop} instead of
linear drop as in Poiesuille driven flows.

The goal of this paper is to give the answer as to how to choose parameters and
resolution to obtain the Bretherton/Taylor problem film thickness for the
diffusion interface model such as the free-energy model \cite{swift}.

>>>>

\title{The binary liquid model for gravity-driven microchannel simulations}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
The motion towards exploration of nano- and microscale phenomena brings new
challenges to simulation tools. The continuous physics approach is limited on
those scales, therefore such methods as Molecular Dynamics (MD) or Lattice
Boltzmann Method (LBM) become appropriate. This work has aim to
reproduce the gravity driven microchannel flow of free surface liquid with the
help of the binary liquid lattice Boltzmann method. In the low capillary number
$Ca$ region the flow is thoroughly described by Bretherton \cite{bretherton},
where the deposition film thickness is proportional to $Ca^{2/3}$. The binary
liquid model is continuous interface multiphase model. Therefore, the issue of
how to properly resolve the interface to obtain proper correlations holds. In
the work we address the binary liquid parameters, as mobility $M$, surface
tension parameter $k$, the free energy constant $A$, grid resolution, to
properly simulate microchannel flow. The work is valuable for industry and
researchers willing to simulate Bretherton phenomena.
\end{abstract}

\section{Introduction}
Taylor/Bretherton \cite{bretherton} flow describes flow of long bubbles in
microchannels. It was found that the front meniscus film width is proportional
to $Ca^{2/3}$ for Capillary numbers less than $0.003$. While there are number
of works which simulate the Bretherton problem with the boundary value method
\cite{ingham-plates,heil-planar}, the simulation of Bretherton problem is of
difficulty for diffusion interface methods, where the interface is spread over
the few grid nodes. In this case the question of proper film width resolution
arises. While the discontinuous interface models are good to simulate
non-complicated geometries phenomena, the continuous interface methods are more
flexible to resolve problem evolving coalescence or droplet breakup. This work
concentrates on the resolving parameters range of the binary liquid lattice
Boltzmann method to validate it across the Bretherton problem.

Lattice Boltzmann method has emerged as successful method to simulate various
phenomena as hydrodynamics \cite{yu}, thermal-flows
\cite{karlin-minimalmodels}, micro-flows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. Based on kinetic nature and as a particle
method LBM tackles relatively easy complex geometries and the incorporation of
physics phenomena on the microscopic level, as multiphase models. Most
multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface in the continuous way. Such representation as it was pointed out
brings issues with the film thickness resolution versus the interface
resolution. In other words, when it comes to the resolution of the film
thickness the diffusion interface should be resolved in such way as to have
negligible effect on physics in the film.

This work examines the binary liquids free-energy LBM model \cite{swift}. The
model simulates binary liquids within the assumption of the uniform overall
density. Therefore, the question of how to simulate the Bretherton problem
where two phases are completely of different densities (liquid and gas) is of
interest. However, the motion of a liquid slug and a bubble in a capillary is
governed by the capillary number. Therefore, the viscosities ratio not the
densities ratio is the critical parameter.

The goal of this paper is to give the answer as to how to choose related to the
binary liquid model parameters, viscosities ratio and grid
resolution to obtain the Bretherton/Taylor problem film thickness for the
diffusion interface model. The paper is organized as follows. We briefly
explain the binary liquid lattice Boltzmann model and summarize literature
review for the Bretherton problem. Then the parameters involved in the
simulation are examined and presented in the results section. The paper is
concluded with main findings.
